%% !TEX root = ../ThesisManuscript_SJ.tex
%%
%%
%% MY PUBLICATIONS
%%_______________________________________________
\chapter*{Scientific production}
\addcontentsline{toc}{chapter}{Scientific production}
\markboth{SCIENTIFIC PRODUCTION}{}

\section*{Peer-reviewed articles}
\label{MyArticles}

\begin{itemize}
	\item[] \underline{Jijón, S.}, Supervie, V., and Breban, R. (2017). Prevention of treatable infectious diseases: a game-theoretic approach. {\it Vaccine}, 37(40):5339--5345. \\doi: \href{https://www.sciencedirect.com/science/article/pii/S0264410X17311283}{10.1016/j.vaccine.2017.08.040}
	
%	\item[] \underline{Jijón, S.}, Supervie, V., and Breban, R. (2020). Can HIV epidemics among men who have sex with men be eliminated through participation to PrEP rollouts? {\it AIDS}.
\end{itemize}

\section*{Submitted articles}
\label{MyArticles}

\begin{itemize}
	\item[] \underline{Jijón, S.}, Supervie, V., and Breban, R. (2020). Can HIV epidemics among men who have sex with men be eliminated through participation to PrEP rollouts? {\it [Submitted to AIDS in January 2021; currently under peer-review]}.
\end{itemize}

%%% Using multibib
%\bibliographystyleSJ{TheseBibSty}
%\bibliographySJ{BibFiles/Bib_These.bib}
%
%\vspace{-185pt}
%\section*{Peer-reviewed articles}
%\vspace{180pt}

\section*{Other scientific contributions}

\subsection*{Oral communications}
\begin{itemize}
	\item[] \underline{Jijón, S.}, Molina, J.-M., Costagliola, D., Supervie, V., and Breban, R. Can HIV epidemics be eliminated through voluntary participation to PrEP rollouts?. EACS 2019 (Basel, Switzerland), November 6--9, 2019.
	
	Abstract in EACS 2019 -- Abstract book. \emph{HIV Medicine}, 20(S9):35.  DOI: 10.1111/ hiv.12814.\\
	\item[] \underline{Jijón, S.}, Prevención de enfermedades infecciosas en el contexto del tratamiento eficiente: un acercamiento por la teoría de juegos y una aplicación al VIH. {\it [Prevention of infectious diseases in the context of efficient treatment: a game-theoretic approach and an application to HIV epidemic]}. 
Seminar organized by the MODEMAT Mathematical Modeling Centre (Quito, Ecuador), February 23, 2017.\\
	\item[] \underline{Jijón, S.}  Can you prevent an epidemic by getting vaccinated?. Pint of Science Festival (Paris), May 21, 2019. [Vulgarization of science].%Website: \url{https://pintofscience.fr/event/its-contagious-tackling-the-dangerous-english}.
\end{itemize}

%\subsection*{Abstract} 
%\begin{itemize}
%	\item[] \underline{Jijón, S.}, Molina, J. M., Costagliola, D., Supervie, V., and Breban, R. (2019). Can HIV epidemics be eliminated using PrEP?. In EACS 2019 -- Abstract book. \emph{HIV Medicine}, 20(S9):35.  DOI: 10.1111/ hiv.12814.
%\end{itemize}

\subsection*{Poster communications}
\begin{itemize}
	\item[] \underline{Jijón, S.}, Molina, J.-M., Costagliola, D., Supervie, V., and Breban, R. Can HIV epidemics be eliminated through voluntary participation to PrEP rollouts?. Presented at ANRS seminar, November 25--26, 2019.\\
	\item[] \underline{Jijón, S.}, Supervie, V., and Breban, R. Prévention des maladies infectieuses : une approche par la théorie des jeux. Presented at the Université des Jeunes Chercheurs organized by Sidaction (Carry-le-Rouet, France), October 14--20, 2017, and at the ED~393 seminar (Saint-Malo, France), October 23--25, 2017.
	\\Winner of the ED~393 Posters awards.\\
	\item[] \underline{Jijón, S.}, Supervie, V., and Breban, R. Prevention of infectious diseases in the context of efficient treatment: a game-theoretic approach. Partial results presented at the ED~393 seminar (Saint-Malo, France), October 24--26, 2016.\\% and the EHESP Interdisciplinary Doctoral Network seminar, Rennes, France on March, 2017.
	\item[] \underline{Jijón, S.}, Supervie, V., and Breban, R. Prévention dans le contexte de traitement efficace~: quel sera l'impact de la prophylaxie pré-exposition sur l'épidémie du VIH~? Presented at the ED~393 seminar (Saint-Malo, France), October 19--21, 2015.
\end{itemize}

%\subsection*{Other scientific communication}
%\begin{itemize}
%	\item[] \underline{Jijón, S.}  Can you prevent an epidemic by getting vaccinated?. Pint of Science Festival (Paris), May 21, 2019. %Website: \url{https://pintofscience.fr/event/its-contagious-tackling-the-dangerous-english}.
%\end{itemize}
%



